Optimal (Multiway) Spatial Joins

### 2024-05-19 msj.tex
yufei

From the perspective of theory, SJs are best understood when $k = 2$, i.e., the {\em pairwise} scenario, where the problem can be solved in $O(n \log n + \out)$ time with planesweep \cite{bcko08}. However, the problem becomes significantly more challenging for $k \ge 3$, i.e., the {\em multiway} scenario. All the solutions developed (see \cite{mp98,mp01,pmt99,???} and the references therein) before 2022 suffer from a worst-case time complexity of $O(n^k)$, offering essentially no improvement over the naive method that simply enumerates the entire cartesian product $R_1 \times R_2 \times ... \times R_k$.

@@

Year 2022 witnessed two independent works \cite{ty22,kcko22} that, although not tackling $k$-SJ directly, imply provably fast $k$-SJ algorithms. Specifically, in \cite{ty22}, Tao and Yi studied several variants of ``interval intersection joins'' under updates. Most relevant to our context is the variant where the input includes, for each $i \in [k]$, a set $\I_i$ of 1D intervals in $\real$, and the join result comprises all $k$-tuples $(I_1, I_2, ..., I_k) \in \I_1 \times \I_2 \times ... \times \I_k$ with $\bigcap_{i=1}^k I_i \neq \emptyset$. The objective is to design a data structure, which, given the insertion (resp., deletion) of an interval in one of the $k$ sets, can identify all the newly-appearing (resp., disappearing) $k$-tuples in the join result in $O((1+\Delta) \cdot \polylog n)$ time, where $n = \sum_{i=1}^k |\I_i|$ and $\Delta$ is the number of such $k$-tuples. Tao and Yi \cite{ty22} presented a structure of $O(n \polylog n)$ space achieving the purpose. Combining their structure with planesweep, one can obtain an algorithm for solving the $k$-SJ problem in $O((n + \out) \cdot \polylog n)$ time.

@@

In \cite{kcko22}, Khamis et al.\ investigated

>>>>

From the perspective of theory, SJs are best understood when $k = 2$, i.e., the {\em pairwise} scenario, where it is folklore that the problem can be solved in $O(n \log n + \out)$ time with planesweep \cite{bcko08}. However, the problem becomes significantly more challenging for $k \ge 3$, i.e., the {\em multiway} scenario. All the solutions developed (see \cite{mp98,mp01,pmt99,???} and the references therein) before 2022 suffer from a worst-case time complexity of $O(n^k)$, offering essentially no improvement over the naive method that simply enumerates the entire cartesian product $R_1 \times R_2 \times ... \times R_k$.

@@

Year 2022 witnessed two independent works \cite{ty22,kcko22} that, although not tackling $k$-SJ directly, imply provably fast $k$-SJ algorithms. Specifically, in \cite{ty22}, Tao and Yi studied several variants of ``interval intersection joins'' under updates. Most relevant to our context is the variant where the input includes, for each $i \in [k]$, a set $\I_i$ of 1D intervals in $\real$, and the join result comprises all $k$-tuples $(I_1,$ $I_2,$ $..., I_k) \in \I_1 \times \I_2 \times ... \times \I_k$ with $\bigcap_{i=1}^k I_i \neq \emptyset$. The objective is to design a data structure, which, given the insertion (resp., deletion) of an interval in one of the $k$ sets, can identify all the newly-appearing (resp., disappearing) $k$-tuples in the join result in $O((1+\Delta) \cdot \polylog n)$ time, where $n = \sum_{i=1}^k |\I_i|$ and $\Delta$ is the number of such $k$-tuples. Tao and Yi \cite{ty22} presented a structure of $O(n \polylog n)$ space achieving the purpose. Combining their structure with planesweep, one can obtain an algorithm for solving the $k$-SJ problem in $O((n + \out) \cdot \polylog n)$ time.

@@

In \cite{kcko22}, Khamis et al.\ investigated a type of joins that extend equi-joins in two ways. First, each attribute value in a relation is an interval (rather than a real value); second, each equality predicate in equi-joins is replaced with an ``non-empty intersection'' predicate on the attributes involved. The $k$-SJ problem can be regarded as a join defined next under the framework of \cite{kcko22}. For each $i \in [k]$, define $R_i$ as a relation over two attributes $X$ and $Y$. For each tuple $\bm{u} \in R_i$, its values $\bm{u}(X)$ and $\bm{u}(Y)$ on the two attributes are both intervals (effectively defining a rectangle). The objective is to output all $(\bm{u}_1, \bm{u}_2, ..., \bm{u}_k) \in R_1 \times R_2 \times ... \times R_k$ such that $\cap_{i=1}^k \bm{u}_i(X) \ne \emptyset$ and $\cap_{i=1}^k \bm{u}_i(Y) \ne \emptyset$.
